### 2019-07-27 Writeup/PERT_Draft_DPM_7_27.tex
Changed test_alt for nova to run

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{amssymb}
\usepackage{accents}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage{framed}
\usepackage[square,sort,comma,numbers]{natbib}
\bibpunct[, ]{(}{)}{,}{a}{}{,}%
\def\bibhang{24pt}%
\def\newblock{\ }%
\def\BIBand{and}
\usepackage{url}
\usepackage{float, subfig}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage{amsthm}
\usepackage{grffile}
\usepackage{sidecap}
\usepackage{pbox}
\usepackage{algorithm}
\usepackage{longtable}
\usepackage[noend]{algpseudocode}

@@

	\section{Introduction} \label{sec:introduction}
	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath}; see \citet{soderlund2004building} for an overview. A project is a collection of activities, between which there are precedence relationships due to logical or technological considerations. A precedence relationship is usually reflected as the start time of one activity following the completion of another. 
	%a required separation time between the start time of two activities. 
	Typically, multiple activities can be processed at the same time, and there is no limit on how many activities can be processed simultaneously, as long as the precedence requirements are satisfied. See~\citet{Elmaghraby77} for a detailed treatment of activity networks. In this setting, ``crashing" is an action that consumes a certain amount of one or more resources and shortens the duration of an activity accordingly~\citep{kuhl2008dynamic}. A deterministic optimization model for crashing an activity network was proposed in the 1960s \citep{fulkerson1961network, kelley1961criticalpath}, in which the goal is to complete the project in minimum time by allocating resources under one or more budget constraints.\\
	When the program evaluation and review technique (PERT) was introduced~\citep{malcolm1959application}, activity durations were modeled as independent beta random variables, and the project duration approximated by a normal distribution. Extensions that allow for more general assumptions followed \citep{Elmaghraby77}, and Monte Carlo simulation plays a role in estimating the expected project span, which is difficult to express analytically \citep{burt1971conditional,van1963letter}.
	Heuristics and simulation-based algorithms have been developed to approximately solve the stochastic project crashing problem \citep{aghaie2009ant, bowman1994stochastic, ke2014genetic, kim2007heuristic}. Another approach to handle uncertainty in activity duration is robust optimization, in which the objective is to minimize the worst case project span over a specified uncertainty set. While affinely adaptive recourse decisions are computationally tractable as linear, or second-order cone, programs, this restriction may lead to suboptimal solutions \citep{chen2008linear, cohen2007stochastic}. However, once recourse decisions can take general form, the robust model is only tractable with rectangular uncertainty sets
	\citep{wiesemann2012robust}. \citet{ahipasaoglu2016distributionally} propose a distributionally robust optimization scheme applied to a PERT network, which reformulates the problem as a semidefinite program or a copositive program, depending on the description of uncertainty. The project crashing optimization problem finds application in project management~\citep{demeulemeester2006project, jaselskis1991allocation,tonchia2018industrial}, machine scheduling~\citep{blazewicz1983scheduling,hall1996machine}, health services scheduling~\citep{cardoen2010operating}, chemical processes~\citep{li2008process}, and digital circuit sizing~\citep{kim2007heuristic}.\\
	In this paper, we propose using the concept of stochastic disruptions to model uncertainty in the duration of activities, which differs from existing approaches in both stochastic programming and robust optimization. A stochastic disruption is an event that may occur at any point in the problem's time horizon and results in a change---typically a significant change---in the system's parameters. A few authors apply this idea in models with discrete time periods, in which the disruption can only occur in a set of specified time periods. \citet{yu2004disruptionmgt} introduce scenario-based optimization models for airline scheduling. \citet{morton2009sealift} introduce a sealift scheduling problem under a finite number of stochastic disruptions within a stochastic programming structure; this model structure ``falls between standard two-stage and multi-stage stochastic programs for a multi-period problem" and reduces the size of the problem (scenario tree) to quadratic, rather than exponential, growth in the number of time periods. Our setting inherits the philosophy of \citet{morton2009sealift}, but enhances the model by allowing the random disruption time to be continuous in the context of an activity network, instead of a prespecified set of fixed time periods. \\

@@

	We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), and it decreases the corresponding durations by \(D_{ik}e_{ij},\ \forall (i,k) \in \cA\), where \(e_{ij} \in [0,1]\) denotes the unit effectiveness of crashing option \(j\). For example, suppose the duration between the start time of activity \(1\) and \(2\) is \(D_{12} = 10\), and applying one unit of crashing option \(1\) to activity \(1\) decreases the duration by half; i.e., \(e_{11} = 0.5\). If we apply \(0.4\) unit of crashing option \(1\) to activity \(1\), \(x_{11} = 0.4\), the required separation between activity \(1\) and \(2\) becomes \(10(1 - 0.4 \cdot 0.5) = 8\). The total cost of crashing cannot exceed a given budget, \(B\). The objective is then to minimize the start time of activity \(T\), and thus, we formulate the deterministic project crashing problem as:
	\begin{subequations} \label{prob:static}
		\min \quad & t_T &\\
		\text{s.t.} \quad &  t_k - t_i \geq D_{ik}\left(1 - \sum_{j \in J_i} e_{ij} x_{ij} \right) \qquad \qquad \forall \,(i,k) \in \cA \label{cons:dSep}\\
		& \sum_{i \in I} \sum_{j \in J_i} b_{ij}x_{ij} \leq B  \label{cons:dBudget}\\
		& \sum_{j \in J_i} x_{ij} \leq 1  \qquad \qquad \forall \,i \in I \label{cons:dSingleBudget}\\
		& 0 \le x_{ij} \leq 1  \qquad \qquad \forall \,i \in I, j \in J_i \label{cons:dxub}\\
		& t_i \geq 0 \qquad \qquad \qquad \;\, \forall\, i \in I. \label{cons:dNonnegt}
	In this formulation, \(t_i\) represents the start time of activity \(i \in I\). We aim to minimize the project span, which is the start time of the terminal activity, \(t_T\). Constraint~\eqref{cons:dSep} guarantees the precedence relationship: if activity \(i\) precedes activity \(k\), activity \(k\) cannot start until time \(t_i + D_{ik} (1- \sum_{j \in J_i} e_{ij} x_{ij})\). Constraint~\eqref{cons:dBudget} is the budget constraint and constraint~\eqref{cons:dSingleBudget} ensures that no more than one unit of crashing option can be applied to an activity. Constraint~\eqref{cons:dNonnegt} enforces nonnegativity for the start time of all activities.\\
	For a project crashing problem under stochastic disruptions, we assume at most one stochastic disruption can occur at a random time in the project span. While this assumption may be limiting in some settings, it is appropriate when it is unlikely for two or more disruptions to occur during the time horizon, and can apply, e.g., for natural disasters, major market crashes, cyber attacks, and work stoppages. For example, suppose we manage a construction project, and we aim to plan against the potential hazard caused by an earthquake or an employee strike. It may be unlikely for two major earthquakes or strikes to affect the same project within the relevant time period. We further assume that for each activity \(i \in I\), the crashing decision needs to be made prior to the start of that activity, which is reasonable, e.g., when contracts are involved in commitment of resources~\citep{oberlender1993project}. We assume a disruption does not affect activities that have already started (including those already finished) at the time of the disruption, but the disruption changes the length of activities that have not yet started according to a known probability distribution. It is usually hard to compute the recourse function directly when random parameters have a continuous distribution, and therefore we use sample average approximation (SAA)~\citep{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, which we denote \(\xi^\omega\), consists of the timing of the disruption, \(H^\omega\), and the magnitude of the disruption via increases in the duration parameters, \(d^\omega_{ik}, \forall (i,k) \in \cA\). \\
	Because we assume at most one disruption, we can model the problem as a two-stage stochastic mixed integer program, in which the timing of the second stage is random. That is, the definition of our stages differs from the usual stochastic programming setting. Here the first stage contains decisions through completion of the project, and we follow this policy if no disruption occurs. And, the second stage characterizes the decisions for each realization of the disruption, which commence at the random time, $H^\omega$. The first stage decision variables are carried out until the disruption if it ever occurs, and after the disruption, the scenario-specific recourse decisions are executed.\\
	The extensive formulation of the two-stage stochastic program is shown as formulation~\eqref{prob:extensive}:
	\begin{subequations} \label{prob:extensive}
		z^* = \min \quad & p^0 t_T + \sum_{\omega \in \Omega} p^\omega t_T^\omega \\
		\text{s.t.} \quad & t_k - t_i \geq D_{ik} \left ( 1 - \sum_{j \in J_i} e_{ij} x_{ij} \right ) \qquad \qquad \forall \, (i,k) \in \cA \label{cons:Sep}\\
		& \sum_{i \in I} \sum_{j \in J_i} b_{ij}x_{ij} \leq B  \label{cons:Budget}\\
		& \sum_{j \in J_i} x_{ij} \leq 1  \qquad \qquad \forall \,i \in I \label{cons:SingleBudget}\\
		& H^\omega + G_i^\omega M \geq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:G1}\\
		& H^\omega - (1 - G_i^\omega) M \leq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:G2}\\
		& t_i^\omega + G_i^\omega M_t \geq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:tG1}\\
		& t_i^\omega - G_i^\omega M_t \leq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:tG2}\\
		& x_{ij}^\omega + G_i^\omega \geq x_{ij} \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:xG1}\\
		& x_{ij}^\omega - G_i^\omega \leq x_{ij} \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:xG2}\\
		& t_k^\omega - t_i^\omega \geq D_{ik} + d_{ik}^\omega G_i^\omega -\sum_{j \in J_i} D_{ik} e_{ij} x_{ij}^\omega - \sum_{j \in J_i} d_{ik}^\omega e_{ij} z_{ij}^\omega \qquad  \forall \,(i,k) \in \cA, \omega \in \Omega \label{cons:scenSep}\\
		& \sum_{i \in I}\sum_{j \in J_i} b_{ij}x_{ij}^\omega \leq B \qquad \qquad \forall \,\omega \in \Omega \label{cons:scenBudget}\\
		& \sum_{j \in J_i} x_{ij}^\omega \leq 1 \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:scenBudget1}\\
		& z_{ij}^\omega \leq G_i^\omega \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:linearize1}\\
		& z_{ij}^\omega \leq x_{ij}^\omega \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:linearize2}\\
		& z_{ij}^\omega \geq G_i^\omega + x_{ij}^\omega - 1 \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:linearize3}\\
		& t_i \geq 0 \qquad \qquad \forall \,i \in I \label{cons:nonnegt}\\
		& t_i^\omega \geq H^\omega G_i^\omega \qquad \qquad \forall\, i \in I, \omega \in \Omega \label{cons:extra} \\
		& 0 \leq x_{ij} \leq 1 \qquad \qquad \forall \,i \in I, j \in J_i\\ 
		& 0 \leq x_{ij}^\omega \leq 1 \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega\\
		& 0 \leq z_{ij}^\omega \leq 1 \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega\\
		& G_i^\omega \in \{0,1\}. \qquad \qquad \forall \,i \in I, \omega \in \Omega. \label{cons:Gbounds}

@@

	\section{NP-Hardness} \label{sec:nphard}
	We show that the optimal project crashing problem under a stochastic disruption is NP-hard even with a single disruption scenario, which occurs with probability one at time zero.
	%The proof process is largely based on the proof of De et al.~1997 for the discrete time-cost tradeoff (\verb|DTCT|) problem. In De et al.~\cite{de1997complexity}, the authors show that an exactly-one-in-three \verb|3SAT| (\verb|EOIT_3SAT|) problem can be transformed to an instance of the \verb|DTCT| decision problem in polynomial time. Since \verb|EOIT_3SAT| is NP-complete (see \cite{Garey1979ComputersAI} for the detailed proof), the authors prove that the \verb|DTCT| decision problem is also NP-complete, and the \verb|DTCT| optimization problem is NP-hard.\\
	%	\newline
	%Similar to De et al.~\cite{de1997complexity}, 
	Our proof relies on a transformation from the exactly-one-in-three \verb|3SAT| (\verb|EOIT_3SAT|) problem: Let \(U = \{u_1,u_2, \dots, u_n\}\) be a set of variables. A literal can be either \(u\) or \(\bar{u} = \neg u\) for \(u \in U\). Let \(C = \{c_1, c_2, \dots, c_m\}\) be a set of clauses, each of which is formed by a disjunction of three literals, e.g., \(c_i = u_j \vee u_k \vee \bar{u}_{\ell}\). The \verb|EOIT_3SAT| problem asks whether there is a truth assignment for each \(u \in U\) such that each clause in \(C\) has exactly one true literal. \citet{de1997complexity} use \verb|EOIT_3SAT| to prove that an activity network problem, in which there are a finite set of alternatives for each activity with different duration and cost, is NP-hard, and we use similar proof constructs. \\

@@

	From now on we refer to the activities representing variables as \(u\)-activities and those representing clauses as \(c\)-activities.\\
	We make the following assignments:
	\begin{subequations}\label{eqn:EOIT_assignments}
		&& D_{u_{jp},c_{ij}}=d^1_{u_{jp},c_{ij}}=0 \qquad \forall i=1,\ldots,m, j=1,\ldots,n, p=2,3: u_j \ \mbox{is in clause} \ c_i \label{eqn:u_c_distance} \\
		&& D_{c_{ij},T}=1 \ \mbox{and} \ d^1_{c_{ij},T}=0 \qquad \forall i=1,\ldots,m, j=1,\ldots,n:  u_j \ \mbox{is in clause} \ c_i  \label{eqn:c_T_distance} \\
		&&e_{c_{ij},1} = 1 \qquad \forall i=1,\ldots,m, j=1,\ldots,n:  u_j \ \mbox{is in clause} \ c_i \label{eqn:e_unit_effectiveness} \\
		&& b_{ij}=1 \qquad \forall i=1,\ldots,m, j=1,\ldots,n:  u_j \ \mbox{is in clause} \ c_i \label{eqn:b_unit_cost} \\
		&& B=2m \label{eqn:budget_is_2m}
	The nominal duration and the disrupted duration for the arcs between \(u\)-activities and \(c\)-activities are \(0\) per equation~\eqref{eqn:u_c_distance}. The nominal and disrupted durations for the arcs between the \(c\)-activities and \(T\) are specified in equation~\eqref{eqn:c_T_distance}. Unlike the \(u\)-activities, each \(c\)-activity can be crashed with a single option with unit effectiveness as given in equation~\eqref{eqn:e_unit_effectiveness}. We assign the budget in equation~\eqref{eqn:budget_is_2m}, where \(m\) is the total number of clauses, and assign unit $b_{ij}$ values in equation~\eqref{eqn:b_unit_cost}.
		\includegraphics[width=0.42\textwidth]{layer2}
		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee \bar{u}_{\ell}\), in the second layer of the constructed activity network for \texttt{EOIT\char`_3SAT} with the arcs connecting it with the first and the third layer}
		\label{fig:layer2}
	\noi We illustrate the logic behind this construction using Figure~\ref{fig:layer2}. For variable \(u_j\), if \(G^1_{u_{j1}} = 1\) then the earliest time activity \(u_{j2}\) can start is \(2\), and activity \(u_{j3}\) can start at time \(1\). If \(G^1_{u_{j1}} = 0\) then the earliest time activity \(u_{j3}\) can start is \(2\), and activity \(u_{j2}\) can start at time \(1\). The same holds for variable \(u_k\), and the opposite for variable \(u_\ell\). The truth assignments indicate which path is longer.
	%\tcr{and which crashing option to select for the \tcr{$c$-activity} nodes}. 
	\\

@@

	\noi
	As a result of Theorem~\ref{thm:npcomplete} we immediately obtain the following result.
	\begin{corollary} \label{thm:nphard}
		Model~\eqref{prob:extensive} is NP-hard, even under a single disruption scenario, which occurs with probability one at time zero.
		%		The project crashing optimization problem under a stochastic disruption is NP-hard.
	\end{corollary}
	%	\begin{proof}[Proof of Corollary~\ref{thm:nphard}]
	%		From Theorem~\ref{thm:npcomplete} states that the project crashing decision problem under a stochastic disruption is NP-complete. This implies the project crashing optimization problem under a stochastic disruption, formulated as model~\ref{prob:extensive}, is NP-hard.
	%	\end{proof}

	\section{Illustration of Problem Properties via Examples} \label{sec:examples}

>>>>

\usepackage{amsmath}
\usepackage{scalefnt}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{color}
\usepackage{amssymb}
\usepackage{accents}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{verbatim}
\usepackage{multirow}
\usepackage{framed}
\usepackage[square,sort,comma,numbers]{natbib}
\bibpunct[, ]{(}{)}{,}{a}{}{,}%
	\def\bibhang{24pt}%
	\def\newblock{\ }%
	\def\BIBand{and}
\usepackage{url}
\usepackage{float, subfig}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage{amsthm}
\usepackage{grffile}
\usepackage{sidecap}
\usepackage{pbox}
\usepackage{algorithm}
\usepackage{longtable}
\usepackage[noend]{algpseudocode}

@@

\section{Introduction} \label{sec:introduction}
	The management of complex projects through optimization has a rich history in operations research, beginning with the critical path method of \cite{kelley1961criticalpath}; see \citet{soderlund2004building} for an overview. A project is a collection of activities, between which there are precedence relationships due to logical or technological considerations. A precedence relationship is usually reflected as the start time of one activity following the completion of another. 
	%a required separation time between the start time of two activities. 
	Typically, multiple activities can be processed at the same time, and there is no limit on how many activities can be processed simultaneously, as long as the precedence requirements are satisfied. See~\citet{Elmaghraby77} for a detailed treatment of activity networks. In this setting, ``crashing" is an action that consumes a certain amount of one or more resources and shortens the duration of an activity accordingly~\citep{kuhl2008dynamic}. A deterministic optimization model for crashing an activity network was proposed in the 1960s \citep{fulkerson1961network, kelley1961criticalpath}, in which the goal is to complete the project in minimum time by allocating resources under one or more budget constraints.

	When the program evaluation and review technique (PERT) was introduced~\citep{malcolm1959application}, activity durations were modeled as independent beta random variables, and the project duration approximated by a normal distribution. Extensions that allow for more general assumptions followed \citep{Elmaghraby77}, and Monte Carlo simulation plays a role in estimating the expected project span, which is difficult to express analytically \citep{burt1971conditional,van1963letter}.
	Heuristics and simulation-based algorithms have been developed to approximately solve the stochastic project crashing problem \citep{aghaie2009ant, bowman1994stochastic, ke2014genetic, kim2007heuristic}. Another approach to handle uncertainty in activity duration is robust optimization, in which the objective is to minimize the worst case project span over a specified uncertainty set. While affinely adaptive recourse decisions are computationally tractable as linear, or second-order cone, programs, this restriction may lead to suboptimal solutions \citep{chen2008linear, cohen2007stochastic}. However, once recourse decisions can take general form, the robust model is only tractable with rectangular uncertainty sets
	\citep{wiesemann2012robust}. \citet{ahipasaoglu2016distributionally} propose a distributionally robust optimization scheme applied to a PERT network, which reformulates the problem as a semidefinite program or a copositive program, depending on the description of uncertainty. The project crashing optimization problem finds application in project management~\citep{demeulemeester2006project, jaselskis1991allocation,tonchia2018industrial}, machine scheduling~\citep{blazewicz1983scheduling,hall1996machine}, health services scheduling~\citep{cardoen2010operating}, chemical processes~\citep{li2008process}, and digital circuit sizing~\citep{kim2007heuristic}.

	In this paper, we propose using the concept of stochastic disruptions to model uncertainty in the duration of activities, which differs from existing approaches in both stochastic programming and robust optimization. A stochastic disruption is an event that may occur at any point in the problem's time horizon and results in a change---typically a significant change---in the system's parameters. A few authors apply this idea in models with discrete time periods, in which the disruption can only occur in a set of specified time periods. \citet{yu2004disruptionmgt} introduce scenario-based optimization models for airline scheduling. \citet{morton2009sealift} introduce a sealift scheduling problem under a finite number of stochastic disruptions within a stochastic programming structure; this model structure ``falls between standard two-stage and multi-stage stochastic programs for a multi-period problem" and reduces the size of the problem (scenario tree) to quadratic, rather than exponential, growth in the number of time periods. Our setting inherits the philosophy of \citet{morton2009sealift}, but enhances the model by allowing the random disruption time to be continuous in the context of an activity network, instead of a prespecified set of fixed time periods.

@@

	We can apply a finite set of crashing options, \(j \in J_i\), to activity \(i \in I\). One unit application of each option incurs a cost of \(b_{ij}\), and it decreases the corresponding durations by \(D_{ik}e_{ij},\ \forall (i,k) \in \cA\), where \(e_{ij} \in [0,1]\) denotes the unit effectiveness of crashing option \(j\). For example, suppose the duration between the start time of activity \(1\) and \(2\) is \(D_{12} = 10\), and applying one unit of crashing option \(1\) to activity \(1\) decreases the duration by half; i.e., \(e_{11} = 0.5\). If we apply \(0.4\) unit of crashing option \(1\) to activity \(1\), \(x_{11} = 0.4\), the required separation between activity \(1\) and \(2\) becomes \(10(1 - 0.4 \cdot 0.5) = 8\). The total cost of crashing cannot exceed a given budget, \(B\). The objective is then to minimize the start time of activity \(T\), and thus, we formulate the deterministic project crashing problem as:
	\begin{subequations} \label{prob:static}
		\begin{align}
		\min \quad & t_T &\\
		\text{s.t.} \quad &  t_k - t_i \geq D_{ik}\left(1 - \sum_{j \in J_i} e_{ij} x_{ij} \right) \qquad \qquad \forall \,(i,k) \in \cA \label{cons:dSep}\\
		& \sum_{i \in I} \sum_{j \in J_i} b_{ij}x_{ij} \leq B  \label{cons:dBudget}\\
		& \sum_{j \in J_i} x_{ij} \leq 1  \qquad \qquad \forall \,i \in I \label{cons:dSingleBudget}\\
		& 0 \le x_{ij} \leq 1  \qquad \qquad \forall \,i \in I, j \in J_i \label{cons:dxub}\\
		& t_i \geq 0 \qquad \qquad \qquad \;\, \forall\, i \in I. \label{cons:dNonnegt}
		\end{align}
	\end{subequations}
	In this formulation, \(t_i\) represents the start time of activity \(i \in I\). We aim to minimize the project span, which is the start time of the terminal activity, \(t_T\). Constraint~\eqref{cons:dSep} guarantees the precedence relationship: if activity \(i\) precedes activity \(k\), activity \(k\) cannot start until time \(t_i + D_{ik} (1- \sum_{j \in J_i} e_{ij} x_{ij})\). Constraint~\eqref{cons:dBudget} is the budget constraint and constraint~\eqref{cons:dSingleBudget} ensures that no more than one unit of crashing option can be applied to an activity. Constraint~\eqref{cons:dNonnegt} enforces nonnegativity for the start time of all activities.

	For a project crashing problem under stochastic disruptions, we assume at most one stochastic disruption can occur at a random time in the project span. While this assumption may be limiting in some settings, it is appropriate when it is unlikely for two or more disruptions to occur during the time horizon, and can apply, e.g., for natural disasters, major market crashes, cyber attacks, and work stoppages. For example, suppose we manage a construction project, and we aim to plan against the potential hazard caused by an earthquake or an employee strike. It may be unlikely for two major earthquakes or strikes to affect the same project within the relevant time period. We further assume that for each activity \(i \in I\), the crashing decision needs to be made prior to the start of that activity, which is reasonable, e.g., when contracts are involved in commitment of resources~\citep{oberlender1993project}. We assume a disruption does not affect activities that have already started (including those already finished) at the time of the disruption, but the disruption changes the length of activities that have not yet started according to a known probability distribution. It is usually hard to compute the recourse function directly when random parameters have a continuous distribution, and therefore we use sample average approximation (SAA)~\citep{kim2015guide,shapiro2009lectures}. In this paper, we assume there is a finite set of scenarios indexed by \(\omega \in \Omega\). For each scenario \(\omega\), the random realization of parameters, which we denote \(\xi^\omega\), consists of the timing of the disruption, \(H^\omega\), and the magnitude of the disruption via increases in the duration parameters, \(d^\omega_{ik}, \forall (i,k) \in \cA\).

	Because we assume at most one disruption, we can model the problem as a two-stage stochastic mixed-integer program, in which the timing of the second stage is random. That is, the definition of our stages differs from the usual stochastic programming setting. Here the first stage contains decisions through completion of the project, and we follow this policy if no disruption occurs. And, the second stage characterizes the decisions for each realization of the disruption, which commence at the random time, $H^\omega$. The first stage decision variables are carried out until the disruption if it ever occurs, and after the disruption, the scenario-specific recourse decisions are executed.

	The extensive formulation of the two-stage stochastic program is shown as formulation~\eqref{prob:extensive}:
	\begin{subequations} \label{prob:extensive}
		\begin{align}
		z^* = \min \quad & p^0 t_T + \sum_{\omega \in \Omega} p^\omega t_T^\omega \\
		\text{s.t.} \quad & t_k - t_i \geq D_{ik} \left ( 1 - \sum_{j \in J_i} e_{ij} x_{ij} \right ) \qquad \qquad \forall \, (i,k) \in \cA \label{cons:Sep}\\
		& \sum_{i \in I} \sum_{j \in J_i} b_{ij}x_{ij} \leq B  \label{cons:Budget}\\
		& \sum_{j \in J_i} x_{ij} \leq 1  \qquad \qquad \forall \,i \in I \label{cons:SingleBudget}\\
		& H^\omega + G_i^\omega M \geq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:G1}\\
		& H^\omega - (1 - G_i^\omega) M \leq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:G2}\\
		& t_i^\omega + G_i^\omega M_t \geq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:tG1}\\
		& t_i^\omega - G_i^\omega M_t \leq t_i \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:tG2}\\
		& x_{ij}^\omega + G_i^\omega \geq x_{ij} \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:xG1}\\
		& x_{ij}^\omega - G_i^\omega \leq x_{ij} \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:xG2}\\
		& t_k^\omega - t_i^\omega \geq D_{ik} + d_{ik}^\omega G_i^\omega -\sum_{j \in J_i} D_{ik} e_{ij} x_{ij}^\omega - \sum_{j \in J_i} d_{ik}^\omega e_{ij} z_{ij}^\omega \qquad  \forall \,(i,k) \in \cA, \omega \in \Omega \label{cons:scenSep}\\
		& \sum_{i \in I}\sum_{j \in J_i} b_{ij}x_{ij}^\omega \leq B \qquad \qquad \forall \,\omega \in \Omega \label{cons:scenBudget}\\
		& \sum_{j \in J_i} x_{ij}^\omega \leq 1 \qquad \qquad \forall \,i \in I, \omega \in \Omega \label{cons:scenBudget1}\\
		& z_{ij}^\omega \leq G_i^\omega \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:linearize1}\\
		& z_{ij}^\omega \leq x_{ij}^\omega \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:linearize2}\\
		& z_{ij}^\omega \geq G_i^\omega + x_{ij}^\omega - 1 \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega \label{cons:linearize3}\\
		& t_i \geq 0 \qquad \qquad \forall \,i \in I \label{cons:nonnegt}\\
		& t_i^\omega \geq H^\omega G_i^\omega \qquad \qquad \forall\, i \in I, \omega \in \Omega \label{cons:extra} \\
		& 0 \leq x_{ij} \leq 1 \qquad \qquad \forall \,i \in I, j \in J_i\\ 
		& 0 \leq x_{ij}^\omega \leq 1 \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega\\
		& 0 \leq z_{ij}^\omega \leq 1 \qquad \qquad \forall \,i \in I, j \in J_i, \omega \in \Omega\\
		& G_i^\omega \in \{0,1\}. \qquad \qquad \forall \,i \in I, \omega \in \Omega. \label{cons:Gbounds}
		\end{align}
	\end{subequations}

@@

\section{NP-Hardness} \label{sec:nphard}
	We show that the optimal project crashing problem under a stochastic disruption is NP-hard even with a single disruption scenario, which occurs with probability one at time zero.
	%The proof process is largely based on the proof of De et al.~1997 for the discrete time-cost tradeoff (\verb|DTCT|) problem. In De et al.~\cite{de1997complexity}, the authors show that an exactly-one-in-three \verb|3SAT| (\verb|EOIT_3SAT|) problem can be transformed to an instance of the \verb|DTCT| decision problem in polynomial time. Since \verb|EOIT_3SAT| is NP-complete (see \cite{Garey1979ComputersAI} for the detailed proof), the authors prove that the \verb|DTCT| decision problem is also NP-complete, and the \verb|DTCT| optimization problem is NP-hard.\\
%	\newline
	%Similar to De et al.~\cite{de1997complexity}, 
	Our proof relies on a transformation from the exactly-one-in-three \verb|3SAT| (\verb|EOIT_3SAT|) problem: Let \(U = \{u_1,u_2, \dots, u_n\}\) be a set of variables. A literal can be either \(u\) or \(\bar{u} = \neg u\) for \(u \in U\). Let \(C = \{c_1, c_2, \dots, c_m\}\) be a set of clauses, each of which is formed by a disjunction of three literals, e.g., \(c_i = u_j \vee u_k \vee \bar{u}_{\ell}\). The \verb|EOIT_3SAT| problem asks whether there is a truth assignment for each \(u \in U\) such that each clause in \(C\) has exactly one true literal. \citet{de1997complexity} use \verb|EOIT_3SAT| to prove that an activity network problem, in which there are a finite set of alternatives for each activity with different duration and cost, is NP-hard, and we use similar proof constructs.

@@

	From now on we refer to the activities representing variables as \(u\)-activities and those representing clauses as \(c\)-activities.

	We make the following assignments:
	\begin{subequations}\label{eqn:EOIT_assignments}
	\begin{eqnarray}
	&& D_{u_{jp},c_{ij}}=d^1_{u_{jp},c_{ij}}=0 \qquad \forall i=1,\ldots,m, j=1,\ldots,n, p=2,3: u_j \ \mbox{is in clause} \ c_i \label{eqn:u_c_distance} \\
	&& D_{c_{ij},T}=1 \ \mbox{and} \ d^1_{c_{ij},T}=0 \qquad \forall i=1,\ldots,m, j=1,\ldots,n:  u_j \ \mbox{is in clause} \ c_i  \label{eqn:c_T_distance} \\
	&&e_{c_{ij},1} = 1 \qquad \forall i=1,\ldots,m, j=1,\ldots,n:  u_j \ \mbox{is in clause} \ c_i \label{eqn:e_unit_effectiveness} \\
	&& b_{ij}=1 \qquad \forall i=1,\ldots,m, j=1,\ldots,n:  u_j \ \mbox{is in clause} \ c_i \label{eqn:b_unit_cost} \\
	&& B=2m \label{eqn:budget_is_2m}
	\end{eqnarray}
	\end{subequations}
     The nominal duration and the disrupted duration for the arcs between \(u\)-activities and \(c\)-activities are \(0\) per equation~\eqref{eqn:u_c_distance}. The nominal and disrupted durations for the arcs between the \(c\)-activities and \(T\) are specified in equation~\eqref{eqn:c_T_distance}. Unlike the \(u\)-activities, each \(c\)-activity can be crashed with a single option with unit effectiveness as given in equation~\eqref{eqn:e_unit_effectiveness}. We assign the budget in equation~\eqref{eqn:budget_is_2m}, where \(m\) is the total number of clauses, and assign unit $b_{ij}$ values in equation~\eqref{eqn:b_unit_cost}.
	\begin{figure}
		\centering
		\includegraphics[width=0.42\textwidth]{layer2}
		\caption{The \(i\)-th clause, \(u_{j} \vee u_{k} \vee \bar{u}_{\ell}\), in the second layer of the constructed activity network for \texttt{EOIT\char`_3SAT} with the arcs connecting it with the first and the third layer}
		\label{fig:layer2}
	\end{figure}
	\noi We illustrate the logic behind this construction using Figure~\ref{fig:layer2}. For variable \(u_j\), if \(G^1_{u_{j1}} = 1\) then the earliest time activity \(u_{j2}\) can start is \(2\), and activity \(u_{j3}\) can start at time \(1\). If \(G^1_{u_{j1}} = 0\) then the earliest time activity \(u_{j3}\) can start is \(2\), and activity \(u_{j2}\) can start at time \(1\). The same holds for variable \(u_k\), and the opposite for variable \(u_\ell\). The truth assignments indicate which path is longer.
	%\tcr{and which crashing option to select for the \tcr{$c$-activity} nodes}.

@@

	As a result of Theorem~\ref{thm:npcomplete} we immediately obtain the following result.
	\begin{corollary} \label{thm:nphard}
		Model~\eqref{prob:extensive} is NP-hard, even under a single disruption scenario, which occurs with probability one at time zero.
%		The project crashing optimization problem under a stochastic disruption is NP-hard.
	\end{corollary}
%	\begin{proof}[Proof of Corollary~\ref{thm:nphard}]
%		From Theorem~\ref{thm:npcomplete} states that the project crashing decision problem under a stochastic disruption is NP-complete. This implies the project crashing optimization problem under a stochastic disruption, formulated as model~\ref{prob:extensive}, is NP-hard.
%	\end{proof}
